Approximation as Differentiation

### 2025-05-05 appendix/notes/auto-diff.tex
Diff is a full subcategory of Fam(FinVect).

Write $\Diff_*$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$, composition in $\Diff_*$ is given by:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m
\linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and
$(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.

@@

>>>>

\begin{definition}[$\Diff$]
\label{def:auto-diff:Diff}
Write $\Diff_*$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$, composition in $\Diff_*$ is given by:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\end{definition}

We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
\RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for any $x \in
\RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp
f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.

@@

If we view any $\Diff$ object $\RR^n$ (\defref{auto-diff:Diff}) as implicitly paired with a constant family
$\{\RR^n\}_{x \in \RR^n}$ and any $\Diff$ morphism $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the
pair of a smooth function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta f_x: \RR^m
\linearto \RR^n\}_{x \in \RR^m}$, then it is clear that $\Diff$ is equivalent to a full subcategory of
$\Fam(\FinVect)$.
